### 2021-02-02 manuscript/manuscript.Rmd
first commit

---
title: Forecasting for emergency medcine
author:
  - name: Author1
    email: emaail@example.com
    affiliation: University1
    footnote: 1
  - name: Author2
    email: email2@example.com
    affiliation: University2
    footnote: 2
address:
  - code: University1
    address: adress1
  - code: University2
    address: adress2
footnote:
  - code: 1
    text: "Corresponding Author"

abstract: |
   The Objective of this paper
keywords:  ""
journal: "Which journal?"
date: "`r Sys.Date()`"
geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
#linenumbers: true
numbersections: true
#output: rticles::elsevier_article
header-includes:  
 \usepackage{adjustbox, float,lscape} #use the 'float' package
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    extra_dependencies: ["flafter"]
    keep_tex: true
bibliography: mybibfile.bib
biblio-style: authoryear
#csl: elsevier-without-titles.csl
#biblio-style: abbrevnat
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(patchwork)
library(tsibble)
library(kableExtra)
library(fable)
library(ggthemes)
library(readxl)
```

```{r, include=FALSE, cache=FALSE}

```

# Introduction

# Research background {#lit}

# Experimental design {#design}

# Result and discussion {#result}

# Conclusion {#conclusion}

# References {#references .unnumbered}