### 2022-10-14 rscript/result/analysing_results.Rmd
Moved R scripts into rscript folder

---
title: "analysing results"
author: "Bahman Rostami-Tabar"
date: "12/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r read-result, options}

hts_result_ets <- read_rds("result/ets_hts_rmsse_long_h.rds")
hts_result_arima <- read_rds("result/arima_hts_rmsse_long_h.rds")
result <- hts_result_ets %>% bind_rows(hts_result_arima)
rm <- result %>% filter(!is.nan(error2_scale))
rmse_for_analysis <- rm %>% filter(!is.infinite(error2_scale))

sum(is.infinite(rm$error2_scale))
```

```{r label, options}

rmsse_approach_level_id <-  rmse_for_analysis %>% group_by(level,approach,.id) %>% 
  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

rmsse_approach_id <-  rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(rmsse=mean(rmsse), .groups = "drop")
```

# overall

```{r label, options}
r <- rmsse_approach_level_id %>% group_by(approach) %>% 
  summarise(mean_rmsse=mean(rmsse), median_rmsse=median(rmsse), .groups = "drop")
r

```

## Total

```{r label, options}
rmsse_total <- rmse_for_analysis %>% filter(level=="Total") %>% dplyr::select(-level) %>% 
  group_by(approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_total, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  ggthemes::theme_few()
```

## control level

```{r label, options}
rmsse_control <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_control_h <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Health board level

```{r label, options}
rmsse_hb <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_hb_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Priority

```{r label, options}
rmsse_priority <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_priority_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Nature of incidents

```{r label, options}
rmsse_nature <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

rmsse_nature_all <- rmsse_nature %>% group_by(approach,level) %>% summarise(rmsse=mean(rmsse), .groups = "drop") 
rmsse_nature_all %>% pivot_wider(names_from = approach, values_from = rmsse)%>% arrange(arima_bu) %>% View()
```

```{r label, options}
rmsse_nature_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

##

```{r label, options}
rr <- rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(mean_rmsse=mean(rmsse),.groups = "drop")
ggplot(rmsse_approach_id)+
  geom_boxplot(aes(y=rmsse, x=approach))+
  coord_cartesian(ylim = c(0,2))
```
##

```{r label, options}
rr <- rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(mean_rmsse=mean(rmsse),.groups = "drop")
ggplot(rmsse_approach_id)+
  geom_boxplot(aes(y=rmsse, x=approach))+
  coord_cartesian(ylim = c(0,2))
```

##

```{r label, options}
rmsse_hd <-  rmse_for_analysis %>% group_by(h,.id,approach) %>% 
  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
ggplot(rmsse_hd)+
  geom_boxplot(aes(y=rmsse, x=factor(h)))+
  facet_wrap(vars(approach))
```