### 2022-11-17 manuscript/01-introduction.qmd
Updated intro

# Introduction {#sec-intro}

A failure to match available resources to demand in Emergency Medical Services (EMS) results in patient flow problems, with serious consequences for patients, staff and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197]. Demand forecasting in EMS helps service planners to avoid the mismatch, potentially providing massive savings in costs and lives, and leading to better patient outcomes. Accurate daily demand forecasting enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, and redeploy staff and vehicles as necessary.

Demand forecasts at EMS are typically required at multiple levels of an organization to inform various planning and decision-making processes [@hulshof2012taxonomic]. There are some planning process at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment. Demand forecasts might also be required at different levels for a specific area of interest such as the nature of demand or the priority level. Moreover, the time series data in EMS has an inherent hierarchical and grouped structure to support such forecasting requirements. Demand for emergency medical services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, and stations/hospitals, or divided into groups such as the nature of incidents or demand priority. Therefore, using forecasting methodologies that account for hierarchical and/or grouped structures of time series in EMS is a natural fit.

However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent forecasts at a single level. Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making. With hierarchical forecasting, plans at any level are based on coherent forecasts and therefore can be aligned. Implementing and sustaining improvements in EMS require alignments and coordination between different stakeholders, without which teams operate in isolation leading to conflicts, duplication work, rework, or work that runs counter to the overall goal to improve the quality of delivery service. Hierarchical forecasting framework can be used as a tool to improve coordination between teams across the care services at the national, sub-national, regional and local levels. The hierarchical forecasting approaches not only create consistent forecasts, but are usually also more accurate than the independent (base) forecasts [@hyndman2011optimal]. To our knowledge, there has been no previous research involving hierarchical and grouped forecasting in the entire field of forecasting for healthcare management.

In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of verified incidents from 2015 to 2020 in a major ambulance service in Great Britain. The data has hierarchical and grouped structures, with hierarchies at the national, control (i.e. sub-national), health board (i.e. regional) levels, as well as groups by priority and nature of incidents. We produce consistent point forecasts and forecast distributions for all levels, which is critical for an effective planning and associated risk management. We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches. We first generate independent/base forecasts using Exponential Smoothing State Space (ETS), Generalized Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method, followed by applying bottom-up and optimal reconciliation approaches. Forecast performance is assessed by the Mean Squared Scaled Error (MSSE) for point forecasts and Continuous Ranked Probability Scores (CRPS) for the probabilistic forecasts.
This paper complies with reproducibility principles [@stodden2013best; @boylan2015reproducibility], and can be applied to any healthcare service (e.g., emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure.

The remainder of this article is structured as follows: In @sec-lit, we provide a brief review of the literature and discuss its limitation to position our work; in @sec-experiment, we present the experiment design describing the data set, forecasting methods and forecast evaluation metrics. In @sec-htc, we discuss the hierarchical time series forecasting approaches to generate both point and probabilistic forecasts. In @sec-results, we present and discuss our results; in @sec-conclusion, we summarize our findings and present ideas for future research.