### 2022-11-18 letter/cover_letter.qmd
add cover letetr and title page

---
author:
  - name: Bahman Rostami Tabar & Rob J. Hyndman 
    affiliation:
      - name: Bahman Rostami-Tabar
        department: Cardiff Business School
        address: 3, Colum Drive
        city: Cardiff
        country: United Kingdom
        postal-code: CF10 3EU
address:
  - Prof. Donald M. Yealy 
  - Editor-in-Chief
  - Annals of Emergency Medicine
subject: Paper submission for peer review
opening: "Dear Prof. Yealy ,"
closing: "Yours Faithfully,"
format: letter-pdf
---

We would be grateful if you could consider for publication our manuscript entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" as an Original Research article in Annals of Emergency Medicine.

Time series forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures. For example the ambulance demand (e.g. arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents, .

In this paper, we aim at filling various gaps identified in the literature, and our contributions to the literature are summarised as follows

Thank you very much for your consideration of this manuscript. We are looking forward to hearing from you in due course.