### 2022-11-18 manuscript/title_page.qmd

---
title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations:
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet. Such demand forecasts are required at national, regional and sub-regional levels, and must take account of the nature of incidents and their priorities. These forecasts are often generated independently by different teams within the organization. As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service. To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation. The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area. We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model. We validate the forecasting approach using time-series cross-validation.
keywords:
  - healthcare
  - emergency services
  - forecast reconciliation
  - ambulance demand
  - regression
execute:
  echo: false
  message: false
  warning: false
  cache: true
pdf-engine: pdflatex
format:
  elsevier-pdf:
    keep-tex: true
    include-in-header: preamble.tex
    journal:
      name: Annals of Emergency Medicine
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

- **Funding information:** A funding awarded by Cardiff Business School allowed BRT to travel to Melbourne to visit RJH to discuss the research design, run the experiment and write the article.

- **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment. BRT obtained a funding to meet RJH at Monash University in Melbourne. BRT and RJH provided the analysis of the time series data. RJH developed forecasting models and hierarchical methods; BRT and RJH use R programming to conduct analysis. BRT drafted the manuscript, and RJH contributed substantially to its revision. BRT takes responsibility for the paper as a whole.