### 2022-11-22 manuscript/06-conclusion.qmd

# Conclusion {#sec-conclusion}

Forecasting problems at Emergency Medical Services often have inherent hierarchical and grouped structures. For example, looking at time series of arrival calls in a clinical desk service, Emergency Department admissions, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest. Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the postcode area. Alternative structures may arise when attributes of interest are crossed rather than nested. For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by the nature of incidents. It is also natural to have a mixed structure, for example, the total demand could be disaggregated by priority and by health board.

Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecasts for each time series independently. This practice may lead to a lack of coordination and possibly undesirable and conflicting outcomes. Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks. In this paper, we investigate the application of hierarchical forecasting methods for producing probabilistic forecasts of daily incidents attended up to 84 days ahead, using different forecasting methods.

Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to the planning and decision-making through producing coherent forecasts, but also it can result in forecast accuracy improvements. Our proposed forecasting models, combined with reconciliation approaches, outperform the empirical distribution benchmark. We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods. When a point forecast is of interest at the bottom level of the series, we observe that reconciliation may not improve the forecast accuracy if the bottom series are noisy and lack systematic patterns. However, forecast reconciliation may result in more accurate forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e. point forecast). Coherent forecasts are also crucial for informing planning activities, and we demonstrate that the proposed models produce coherent forecasts across all forecast horizons. Therefore, we recommend that forecast reconciliation approaches be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted. Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individually, works remarkably well for our mixed hierarchical & grouped structure.

Further research could investigate the practical benefits of probabilistic hierarchical forecasting in EMS. Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through a more holistic planning approach. While we generated count distributions of attended incidents for the base forecasts using Poisson regression models, the reconciled forecast distributions are not counts. This could be also an avenue for further research.