### 2023-07-18 manuscript/02-literature.qmd

# Research background {#sec-lit}

Emergency medical services (EMS) are a critical component in the delivery of urgent medical care to communities. An effective service delivery requires accurate resource planning that generally relies on demand forecasts at operational, tactical, and strategic levels. There is a substantial number of studies on the application of time series forecasting in the Emergency Medical Services. For example, @ibrahim2016modeling provide an extensive review of the models used in forecasting call volume arrivals. Another important area is related to forecasting ambulance demand. Although the definition of demand might not be always clearly stated, this is typically referring to a situation where a physical resource has been deployed to respond to an incident. This might be also called _attended incidents_. Another demand related variable is verified incidents; these are all incidents that require an action: either by sending a physical vehicle, responding via the Clinical Support Desk, requesting an external provider to respond to it, or forwarding it to other channels such as police, firefighters, or general practitioners. Our study is aligned with this stream of literature. Another similar area that has received considerable attention is Emergency Department forecasting; we refer interested readers to @mingliterature2022, @gul2020exhaustive, and @wargon2009systematic for extensive reviews of the relevant literature. In this section, we provide a brief review of studies on forecasting ambulance demand in EMS.

There are generally two main streams of research related to forecasting ambulance demand in EMS: (i) the first stream focuses on the application of time series methods and regression approaches to forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; and (ii) the second stream considers forecasting EMS demand in finer temporal and geographical granularities by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit]. The focus of our study is related to the first stream of research.

@sasaki2010using develop a multivariable regression model to estimate future EMS demands. In addition to the historical demand, the population census for different age groups and counts of the number of companies employing more than five people are included in the regression. The census variables describe groups who are more likely to need an ambulance. A stepwise ordinary least squares regression analysis is used for estimating the parameter and generating forecasts. The only performance measure reported in this study is $R^2$, which is not an effective measure of forecast accuracy [@armstrong01, p457]. The research design of this study is not rigorous and the study is not reproducible. @vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day, and 28-day forecast horizons using data provided by an ambulance service in Great Britain. The performance of this approach is compared to Auto-Regressive Integrated Moving Average (ARIMA) and Holt-Winters time series methods using Root Mean Squared Error (RMSE). They concluded that point forecasts generated by SSA are more accurate for longer-term, but that ARIMA and Holt-Winters performance is superior for shorter-term horizons. @vile2016time further develop a decision support system to integrate forecasts generated by SSA. However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization and response time. The tool contains options that allow generating forecasts at various levels of granularity; however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination. @al2021empirical \textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately. The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand. Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated. The findings indicate that ARIMA performs the best in predicting weekly and monthly demand. However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2). Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days. The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval. The findings indicate that the ME consistently produces the most accurate point forecasts. On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead. For a two-week leading period, MU shows better coverage probability.}

@9659837 investigate forecasting EMS demand in a high Spatio-temporal resolution of 1km$^2$ spatial regions and 1-hr time intervals using total incidents in Oslo, Norway, from 1 January 2015 to 11 February 2019. They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare the results to simple aggregation methods and baselines. The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE), and the forecast distribution is measured by Categorical Cross-Entropy. They found that Neural Network models performed better in producing point forecasts, while a distribution baseline method based on the spatial distribution of the incidents across all time steps provided more accurate forecast distributions. @zhou2016predictinglit proposed three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict hourly data 4 weeks ahead for a 1km$^2$ spatial region. Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia were used to build the model and examine the performance on test data using mean negative log-likelihood. They show that forecasts generated by the proposed methods were significantly more accurate than the current industry practice (a simple averaging formula). @grekousis2019will investigated the combination of spatial analysis methods with data mining techniques based on an improved Hungarian algorithm and a MLP neural network to identify the most likely locations of future emergency events. The proposed approach was tested using data from 2851 events attended by the EMS in Athens, Greece, over 24 weeks. They showed that 23% of real emergency events lie within 50 meters of the predicted ones and nearly 70% of the real emergency events lie no further than 150 meters away, which is rather accurate given the granularity of the problem at the city level.

We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:

1. Current studies ignore the inherent hierarchical and/or grouped structure of the time series data, and the relationship between series at different levels of hierarchy. While the hierarchical forecasting methodology has been developed and applied in various domains over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area.

2. Current research is mainly concerned with generating point forecasts at a single level of hierarchy. There is a lack of studies considering the entire forecast distribution of daily ambulance demand for the whole hierarchy to better represent the uncertainty of future demand, providing a risk management tool for planners.

3. Reproducibility is still a major challenge in EMS forecasting, as it is unlikely that any reader can  reproduce prior studies without the help of the authors of those papers.

4. Another limitation is related to the generated forecasts not being on the sample space of non-negative counts. Since actual ambulance counts cannot be negative or non-integer, ambulance demand forecast distributions should reflect the data. Of course, point forecasts represent means, so they should be non-negative, but may be non-integer. While this might not be an issue when producing forecasts at a single level, producing non-negative count forecasts in a hierarchical/grouped structure is challenging and requires further investigation in the future.

```{r summarylit, echo=FALSE, cache=FALSE}
library(kableExtra)
li <- readxl::read_xlsx("../literature_table/table.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies in forecasting FOR AMBULANCE SERVICES") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
```