### 2023-07-18 manuscript/06-conclusion.qmd

# Conclusion {#sec-conclusion}

Forecasting problems at Emergency Medical Services often have inherent hierarchical and grouped structures. For example, looking at time series of arrival calls in a clinical desk service, Emergency Department admissions, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest. Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the postcode area. Alternative structures may arise when attributes of interest are crossed rather than nested. For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by the nature of incidents. It is also natural to have a mixed structure, for example, the total demand could be disaggregated by priority and by health board.

Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecasts for each time series independently. This practice may lead to a lack of coordination and possibly undesirable and conflicting outcomes. Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks. In this paper, we investigate the application of hierarchical forecasting methods for producing probabilistic forecasts of daily incidents attended up to 84 days ahead, using different forecasting methods.

Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to the planning and decision-making through producing coherent forecasts, but also it can result in forecast accuracy improvements. Our proposed forecasting models, combined with reconciliation approaches, outperform the empirical distribution benchmark. We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods. When a point forecast is of interest at the bottom level of the series, we observe that reconciliation may not improve the forecast accuracy if the bottom series are noisy and lack systematic patterns. However, forecast reconciliation may result in more accurate forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e. point forecast). Coherent forecasts are also crucial for informing planning activities, and we demonstrate that the proposed models produce coherent forecasts across all forecast horizons. Therefore, we recommend that forecast reconciliation approaches be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted. Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individually, works remarkably well for our mixed hierarchical & grouped structure.

\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS. Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy. We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast. While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact. For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary. This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services. Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself. This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs. These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}

\textcolor{teal}{Future research can build upon this study in several ways. In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the decision impact of this framework on utilities that are significant to the EMS. Linking forecasts with its utilities (e.g. response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.} Aditionally, in our study, we employed Poisson regression models to generate count distributions of attended incidents for the base forecasts. However, it is essential to note that the reconciled forecast distributions are not counts. This observation presents an interesting avenue for future research. \textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available. It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}

\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries} [@ostrom2010moving]. \textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors. This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system. For instance, in the supply chains } [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts. Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency. In the financial industry} [@kimes1998strategic], \textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand. Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation. Likewise, in the hospitality and tourism industry} [@dekimpe2016impact], \textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses. Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation. Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}