### 2023-11-10 rscript/result/analysing_results.Rmd

---
title: "analysing results"
author: "Bahman Rostami-Tabar"
date: "12/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r read-result, options}
hts_result_ets <- read_rds("result/ets_hts_rmsse_long_h.rds")
hts_result_arima <- read_rds("result/arima_hts_rmsse_long_h.rds")
result <- hts_result_ets %>% bind_rows(hts_result_arima)
rm <- result %>% filter(!is.nan(error2_scale))
rmse_for_analysis <- rm %>% filter(!is.infinite(error2_scale))

sum(is.infinite(rm$error2_scale))
```

```{r label, options}
rmsse_approach_level_id <- rmse_for_analysis %>%
  group_by(level, approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

rmsse_approach_id <- rmsse_approach_level_id %>%
  group_by(approach, level) %>%
  summarise(rmsse = mean(rmsse), .groups = "drop")
```

# overall

```{r label, options}
r <- rmsse_approach_level_id %>%
  group_by(approach) %>%
  summarise(mean_rmsse = mean(rmsse), median_rmsse = median(rmsse), .groups = "drop")
r
```

## Total

```{r label, options}
rmsse_total <- rmse_for_analysis %>%
  filter(level == "Total") %>%
  dplyr::select(-level) %>%
  group_by(approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_total, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
  geom_boxplot() +
  ggthemes::theme_few()
```

## control level

```{r label, options}
rmsse_control <- rmse_for_analysis %>%
  filter(level %in% c("G1/S", "G1/N", "G1/C")) %>%
  group_by(level, approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
  geom_boxplot() +
  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
  ggthemes::theme_few()
```

```{r label, options}
rmsse_control_h <- rmse_for_analysis %>%
  filter(level %in% c("G1/S", "G1/N", "G1/C")) %>%
  group_by(level, approach, h) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control_h, aes(x = h, y = rmsse, color = factor(approach))) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 35, 1)) +
  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
  ggthemes::theme_few() +
  ggthemes::scale_color_colorblind()
```

## Health board level

```{r label, options}
rmsse_hb <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level, approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
  geom_boxplot() +
  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
  ggthemes::theme_few()
```

```{r label, options}
rmsse_hb_h <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level, approach, h) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x = h, y = rmsse, color = factor(approach))) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 35, 1)) +
  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
  ggthemes::theme_few() +
  ggthemes::scale_color_colorblind()
```

## Priority

```{r label, options}
rmsse_priority <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level, approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
  geom_boxplot() +
  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
  ggthemes::theme_few()
```

```{r label, options}
rmsse_priority_h <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level, approach, h) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority_h, aes(x = h, y = rmsse, color = factor(approach))) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 35, 1)) +
  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
  ggthemes::theme_few() +
  ggthemes::scale_color_colorblind()
```

## Nature of incidents

```{r label, options}
rmsse_nature <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level, approach, .id) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

rmsse_nature_all <- rmsse_nature %>%
  group_by(approach, level) %>%
  summarise(rmsse = mean(rmsse), .groups = "drop")
rmsse_nature_all %>%
  pivot_wider(names_from = approach, values_from = rmsse) %>%
  arrange(arima_bu) %>%
  View()
```

```{r label, options}
rmsse_nature_h <- rmse_for_analysis %>%
  filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level, approach, h) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x = h, y = rmsse, color = factor(approach))) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 35, 1)) +
  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
  ggthemes::theme_few() +
  ggthemes::scale_color_colorblind()
```

##

```{r label, options}
rr <- rmsse_approach_level_id %>%
  group_by(approach, level) %>%
  summarise(mean_rmsse = mean(rmsse), .groups = "drop")
ggplot(rmsse_approach_id) +
  geom_boxplot(aes(y = rmsse, x = approach)) +
  coord_cartesian(ylim = c(0, 2))
```
##

```{r label, options}
rr <- rmsse_approach_level_id %>%
  group_by(approach, level) %>%
  summarise(mean_rmsse = mean(rmsse), .groups = "drop")
ggplot(rmsse_approach_id) +
  geom_boxplot(aes(y = rmsse, x = approach)) +
  coord_cartesian(ylim = c(0, 2))
```

##

```{r label, options}
rmsse_hd <- rmse_for_analysis %>%
  group_by(h, .id, approach) %>%
  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
ggplot(rmsse_hd) +
  geom_boxplot(aes(y = rmsse, x = factor(h))) +
  facet_wrap(vars(approach))
```